---
title: "meta_analysis parental SEP and cardiometabolic and inflammatory biomarkers"
author: "Juan Carlos Rivillas"
date: "02/02/2023"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

Meta package documentation in R
#Install packages require to complete meta-analysis.
```{r}
install.packages("meta")
install.packages("metafor")
install.packages("gridGraphics")
library(metadat)
library(meta)
library(metafor)
library(readxl)

library(meta)
library(vcd)
library(gridGraphics)
library(gridExtra)
```

The meta-analysis are organised as follow:

1) Cardiovascular function
2) Glucose metabolism
3) Lipid metabolism
4) Adiposity
5) Inflammation

I wrote this script to calculate power for a random-effects meta-analysis based on https://journals.sagepub.com/doi/abs/10.3102/1076998609346961 

######### Cardiovascular functions #########

***Blood pressure***
Combinations of the SBP and DBP estimates from cohort studies
```{r}
library(readxl)
datasbp <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/databp_cohort_parentalSEP.xlsx")
View(databp_cohort_parentalSEP)
View(databp_cross_parentalSEP)
```


```{r}
# meta-analysis with continuout outcome
# comb.fixed/comb.random: indicator whether a fix/random effect mata-analysis to be conducted.
# sm: Three different types of summary measures to choose,standardized mean difference (SMD),mean difference (MD), ratio of means (ROM)

#cohort studies and parental SEP (n=13)
res.bp.cohort.parentalSEP =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = databp_cohort_parentalSEP, sm = "SMD")
res.bp.cohort.parentalSEP

#cross-sectional studies and parental SEP (n=7)
res.bp.cross.parentalSEP =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = databp_cross_parentalSEP, sm = "SMD")
res.bp.cross.parentalSEP
```

#Random Forest is an ensemble of decision trees algorithms that can be used for classification and regression predictive modeling.
```{r}
#Random forest regression to create an uncorrelated forest of decision trees.
#cohort studies blood pressure and parental SEP
forest(res.bp.cohort.parentalSEP, comb.fixed = F)

#cross-sectional studies and parental SEP
forest(res.bp.cross.parentalSEP, comb.fixed = F)
```
#Power analysis
#Funnel plots are a visual tool for investigating publication and other bias in meta-analysis. They are simple scatterplots of the treatment effects estimatedfrom individual studies (horizontal axis) against a measure of study size (vertical axis).
```{r}
#Funnel Plot to check or Test for funnel plot asymmetry, based on rank correlation or linear regression method (Metabias).
funnel(res.bp.cohort.parentalSEP)
funnel(res.bp.cross.parentalSEP)
```

Use Egger’s test to check publication bias, can take string ‘Egger’ or ‘linreg’.
```{r}
#We have sufficient number of studies to test for small study effects. 
metabias(res.bp.cohort.parentalSEP, method.bias = 'linreg', k.min = 13, plotit = T)
metabias(res.bp.cross.parentalSEP, method.bias = 'linreg', k.min = 7, plotit = T) 
```

#Investigation of heterogeneity: as we get low p value of Q2 and high I2 we conducted a meta-regression.
```{r}
mu8 <- update(res.bp.cohort.parentalSEP, byvar = Age, tau.common = TRUE, comb.fixed = FALSE)
mu9 <- update(res.bp.cross.parentalSEP, byvar = Age, tau.common = TRUE, comb.fixed = FALSE)
mu8
mu9

metareg(mu8)
metareg(mu9)
#Note: here, contrasts can be applied only to factors with 2 or more levels
# As argument 'tau.common' was used to create mu8

# Generate bubble plot
bubble(metareg(mu8))
#bubble(metareg(mu9)) #1 study with NAs omitted from model fitting then dropped from the model.
```



########### GLUCOSE METABOLISM #############

HbA1c, fasting flucose and blood glucose from cohorts studies.
There were not sufficient cross-sectional studies to perform random forest models for these biomarkers.

Combinations of the SBP and DBP estimates from cohort studies
```{r}
library(readxl)
datasbp <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/dataglucose_cohort_parentalSEP.xlsx")
View(ddataglucose_cohort_parentalSEP)
```


```{r}
# meta-analysis with continuous outcome
# comb.fixed/comb.random: indicator whether a fix/random effect mata-analysis to be conducted.
# sm: Three different types of summary measures to choose,standardized mean difference (SMD),mean difference (MD), ratio of means (ROM)

#cohort studies and parental SEP (n=5)
res.glucose.cohort.parentalSEP =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = dataglucose_cohort_parentalSEP, sm = "SMD")
res.glucose.cohort.parentalSEP
```


```{r}
#cohort studies blood pressure and parental SEP
forest(res.glucose.cohort.parentalSEP, comb.fixed = F)
```
```{r}
#Funnel Plot to check or Test for funnel plot asymmetry, based on rank correlation or linear regression method (Metabias).
funnel(res.glucose.cohort.parentalSEP)

#Use Egger’s test to check publication bias, can take string ‘Egger’ or ‘linreg’.
metabias(res.glucose.cohort.parentalSEP, method.bias = 'linreg', k.min = 5, plotit = T)

#Investigation of heterogeneity: as we get low p value of Q2 and high I2 we conducted a meta-regression.
mu7 <- update(res.glucose.cohort.parentalSEP, byvar = Age, tau.common = TRUE, comb.fixed = FALSE)
metareg(mu7)

mu7

# Generate bubble plot
bubble(metareg(mu7))
```

#######LIPID METABOLISM###########

****Total cholesterol, triglyceride and LDL****
HDL had small number of studies to generate random forest.
```{r}
library(readxl)
datatchol <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/datalipid_cohort_parentalSEP.xlsx")
View(datatlipids_cohort_parentalSEP)
```

```{r}
# meta-analysis with continuous outcomes (n=13)
# comb.fixed/comb.random: indicator whether a fix/random effect meta-analysis to be conducted.
# Type of choose summary measure: standardized mean difference (SMD).
res.lipids.parentalSEP =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = datatlipids_cohort_parentalSEP, sm = "SMD")
res.lipids.parentalSEP

#Random forest regression to create an uncorrelated forest of decision trees algorithms that can be used for classification and regression predictive modeling.
forest(res.lipids.parentalSEP, comb.fixed = F)

#Funnel Plot to check or Test for funnel plot asymmetry, based on rank correlation or linear regression method (Metabias).
funnel(res.lipids.parentalSEP)

#Use Egger’s test to check publication bias, can take string ‘Egger’ or ‘linreg’.
metabias(res.lipids.parentalSEP, method.bias = 'linreg', k.min = 13, plotit = T)

#Investigation of heterogeneity: as we get low p value of Q2 and high I2 we conducted a meta-regression.
mu4 <- update(res.lipids.parentalSEP, byvar = Age, tau.common = TRUE, comb.fixed = FALSE)
metareg(mu4)

mu4

# Generate bubble plot
bubble(metareg(mu4))
```
```


############ADIPOSITY####################

*****BMI and WC****

```{r}
#Upload data cohort studies
library(readxl)
databmi <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/Meta-analysis inflammation & parental SEP/Tables/dataadiposity_cohort_parentalSEP.xlsx")
View(dataadiposity_cohort_parentalSEP)
```


```{r}
# meta-analysis with continuous outcome (n=12)
# comb.fixed/comb.random: indicator whether a fix/random effect meta-analysis to be conducted.
# Type of choose summary measure: standardized mean difference (SMD).
res.adiposity.cohort.parentalSEP =  metacont(n_exposed, mean_exposed, sd_exposed, n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = dataadiposity_cohort_parentalSEP, sm = "SMD")
res.adiposity.cohort.parentalSEP

#Random forest regression to create an uncorrelated forest of decision trees algorithms that can be used for classification and regression predictive modeling.
forest(res.adiposity.cohort.parentalSEP, comb.fixed = F)

#Funnel Plot to check or Test for funnel plot asymmetry, based on rank correlation or linear regression method (Metabias).
funnel(res.adiposity.cohort.parentalSEP)

#Use Egger’s test to check publication bias, can take string ‘Egger’ or ‘linreg’.
metabias(res.adiposity.cohort.parentalSEP, method.bias = 'linreg', k.min = 12, plotit = T) 

#Investigation of heterogeneity: as we get low p value of Q2 and high I2 we conducted a meta-regression.
mu6 <- update(res.adiposity.cohort.parentalSEP, byvar = Sex, tau.common = TRUE, comb.fixed = FALSE)
mu6

# Generate bubble plot
metareg(mu6)
```


##########################INFLAMMATION############################

***CRP and IL-6***

#Upload table with studies CRP biomarkers (cohort studies)
```{r}
setwd("~/Documents/PhD Project/Literature review/systematic review 1 inflammation")

library(readxl)
datacrp_cohorts_parentalSEP <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/datacrp_cohorts_parentalSEP.xlsx")
View(datacrp_cohorts_parentalSEP)

#Check values in the table cohort studies Il-6 & Allostatic Load and parental SEP
View(datainflamm_cohort_parentalSEP)
head(datainflamm_cohort_parentalSEP)

#Check values in the table CRP, IL-6 and AL cross sectional studies and parental SEP
View(datainflamm_cross_parentalSEP)
head(datainflamm_cross_parentalSEP)
```


```{r}
#cohort studies CRP and parental SEP (n=11 studies)
# Meta-analysis with continuout outcome
# comb.fixed/comb.random: indicator whether a fix/random effect mata-analysis to be conducted.
# sm: Three different types of summary measures to choose,standardized mean difference (SMD),mean difference (MD), ratio of means (ROM)

res.crp.cohort.parentalSEP =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = datacrp_cohorts_parentalSEP, sm = "SMD")
res.crp.cohort.parentalSEP

#IL-6 - cohort studies and parental education (n=5 studies)
res.inflamm.cohort.parentalSEP =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = datainflamm_cohort_parentalSEP, sm = "SMD")
res.inflamm.cohort.parentalSEP


#CRP, IL-6 and Allostatic Load - cross sectional studies and parental education (5 studies)
res.inflamm.cross.parentalSEP =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = datainflamm_cross_parentalSEP, sm = "SMD")
res.inflamm.cross.parentalSEP
```


#Random forest regression to create an uncorrelated forest of decision trees.
#Random Forest is an ensemble of decision trees algorithms that can be used for classification and regression predictive modeling.
```{r}
#random forest cohort studies and parental SEP
forest(res.crp.cohort.parentalSEP, comb.fixed = F)

#random forest IL-6 and AL cohort studies and parental SEP
forest(res.inflamm.cohort.parentalSEP, comb.fixed = F)

#random forest CRP, IL-6 and AL cross sectional studies and parental SEP
forest(res.inflamm.cross.parentalSEP, comb.fixed = F)
```
  

```{r}
#Funnel Plot using meta package
funnel(res.crp.cohort.parentalSEP)
funnel(res.inflamm.cohort.parentalSEP)
funnel(res.inflamm.cross.parentalSEP)

#Funnel Plot using metafor package (as a sensitivity analysis)
metafor::funnel(res.crp.cohort.parentalSEP)
metafor::funnel(res.inflamm.cohort.parentalSEP)
metafor::funnel(res.inflamm.cross.parentalSEP)
```


#Use Egger’s test to check publication bias, can take string ‘Egger’ or ‘linreg’.
#Egger’s test is a linear regression of the intervention effect estimates on their standard errors weighted by their inverse variance
```{r}
metabias(res.crp.cohort.parentalSEP, method.bias = 'linreg', k.min = 11, plotit = T)
metabias(res.inflamm.cohort.parentalSEP, method.bias = 'linreg', k.min = 5, plotit = T)
metabias(res.inflamm.cross.parentalSEP, method.bias = 'linreg', k.min = 5, plotit = T)
```


#Investigation of heterogeneity: as we get low p value of Q2 and high I2 we conducted a meta-regression.
```{r}
mu1 <- update(res.crp.cohort.parentalSEP, byvar = Age, tau.common = TRUE, comb.fixed = FALSE)
mu2 <- update(res.inflamm.cohort.parentalSEP, byvar = Age, tau.common = TRUE, comb.fixed = FALSE)
mu3 <- update(res.inflamm.cross.parentalSEP, byvar = Age, tau.common = TRUE, comb.fixed = FALSE)

mu1
mu2
mu3

# Generate bubble plot
bubble(metareg(mu1))
bubble(metareg(mu2))
#bubble(metareg(mu3)) #redundant predictors dropped from the model. 
```



#AL - cohort studies and parental education (n=5 studies)

```{r}
#Allostatic Load - cohort studies and parental education (n=9 studies)
res.al.cohort.parentalSEP =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = dataal_cohort_parentalSEP, sm = "SMD")
res.al.cohort.parentalSEP

#random forest IL-6 and AL cohort studies and parental SEP
forest(res.al.cohort.parentalSEP, comb.fixed = F)

metabias(res.al.cohort.parentalSEP, method.bias = 'linreg', k.min = 4, plotit = T)
mu10 <- update(res.al.cohort.parentalSEP, byvar = Age, tau.common = TRUE, comb.fixed = FALSE)
bubble(metareg(mu10))

```

########END###########

This script are part of appendix of individual meta-analysis (specific biomarkers). Please do not run as a part of the meta-analysis and use it for reaserch or teaching purposes.

#Power analysis (this code was used to extract results to the manuscript)

```{r}

# Download packages

install.packages("devtools")
install.packages("metameta")

library(devtools)
devtools::install_github("dsquintana/metameta")
library(metameta)

#CRP cohort studies and occupation

#effects sizes (means) and standard deviation

View (datacrp_cohort_occupation_pa)        # View the exmaple of dataset

power_crp_cohort_occup <- mapower_se(dat=datacrp_cohort_occupation_pa,  # The dataset
                                     observed_es = 0.07,                # The observed summary effect size 
                                     name = "crp_cohort_occupation_ma") # The name of the meta-analysis

crp_cohort_occupation_table <- power_crp_cohort_occup$dat  # Power for a range of effects
crp_cohort_occupation_table
write.csv (crp_cohort_occupation_table, "crp_cohort_occupation_table.csv") # Save as csv

crp_cohort_occupation_power_med_table <- power_crp_cohort_occup$power_median_dat     # Median power
crp_cohort_occupation_power_med_table
write.csv (crp_cohort_occupation_power_med_table, "crp_cohort_occupation_power_med_table.csv") # Save as csv

firepower (list(power_crp_cohort_occup$power_median_dat))  # Firepower plot

#CRP cohort studies and education

View (datacrp_cohort_education_pa)        # View the exmaple of dataset

power_crp_cohort_education <- mapower_se(dat=datacrp_cohort_education_pa,  # The dataset
                                     observed_es = 0.11,                # The observed summary effect size 
                                     name = "crp_cohort_education_ma") # The name of the meta-analysis

crp_cohort_education_table <- power_crp_cohort_education$dat  # Power for a range of effects
crp_cohort_education_table
write.csv (crp_cohort_education_table, "crp_cohort_education_table.csv") # Save as csv

crp_cohort_education_power_med_table <- power_crp_cohort_education$power_median_dat     # Median power
crp_cohort_education_power_med_table
write.csv (crp_cohort_education_power_med_table, "crp_cohort_education_power_med_table.csv") # Save as csv

firepower (list(power_crp_cohort_education$power_median_dat))  # Firepower plot


#IL-6 cohort studies and education
View (datail6_cohort_education_pa)        # View the exmaple of dataset

power_il6_cohort_education <- mapower_se(dat=datail6_cohort_education_pa,  # The dataset
                                     observed_es = 0.07,                # The observed summary effect size 
                                     name = "il6_cohort_education_ma") # The name of the meta-analysis

il6_cohort_education_table <- power_il6_cohort_education$dat  # Power for a range of effects
il6_cohort_education_table
write.csv (il6_cohort_education_table, "il6_cohort_education_table.csv") # Save as csv

il6_cohort_education_power_med_table <- power_il6_cohort_education$power_median_dat     # Median power
il6_cohort_education_power_med_table
write.csv (il6_cohort_education_power_med_table, "il6_cohort_education_power_med_table.csv") # Save as csv

firepower (list(power_il6_cohort_education$power_median_dat))  # Firepower plot


#SBP cohort studies and education
View (datasbp_cohort_education_pa)        # View the exmaple of dataset

power_sbp_cohort_education <- mapower_se(dat=datasbp_cohort_education_pa,  # The dataset
                                     observed_es = 0.25,                # The observed summary effect size 
                                     name = "sbp_cohort_education_ma")  # The name of the meta-analysis

sbp_cohort_education_table <- power_sbp_cohort_education$dat  # Power for a range of effects
sbp_cohort_education_table
write.csv (sbp_cohort_education_table, "sbp_cohort_education_table.csv") # Save as csv

sbp_cohort_education_power_med_table <- power_sbp_cohort_education$power_median_dat     # Median power
sbp_cohort_education_power_med_table
write.csv (dbp_cohort_education_power_med_table, "sbp_cohort_education_power_med_table.csv") # Save as csv

firepower (list(power_sbp_cohort_education$power_median_dat))  # Firepower plot


#Total cholesterol cohort studies and education
View (datatchol_cohort_education_pa)        # View the exmaple of dataset

power_tchol_cohort_education <- mapower_se(dat=datatchol_cohort_education_pa,  # The dataset
                                     observed_es = 0.19,                # The observed summary effect size 
                                     name = "tchol_cohort_education_ma")  # The name of the meta-analysis

tchol_cohort_education_table <- power_tchol_cohort_education$dat  # Power for a range of effects
tchol_cohort_education_table
write.csv (tchol_cohort_education_table, "tchol_cohort_education_table.csv") # Save as csv

tchol_cohort_education_power_med_table <- power_tchol_cohort_education$power_median_dat     # Median power
tchol_cohort_education_power_med_table
write.csv (tchol_cohort_education_power_med_table, "tchol_cohort_education_power_med_table.csv") # Save as csv

firepower (list(power_tchol_cohort_education$power_median_dat))  # Firepower plot



#LDL cohort studies and education
View (dataldl_cohort_education_pa)        # View the example of dataset

power_ldl_cohort_education <- mapower_se(dat=dataldl_cohort_education_pa,  # The dataset
                                     observed_es = 0.17,                # The observed summary effect size 
                                     name = "ldl_cohort_education_ma")  # The name of the meta-analysis

ldl_cohort_education_table <- power_ldl_cohort_education$dat  # Power for a range of effects
ldl_cohort_education_table
write.csv (ldl_cohort_education_table, "ldl_cohort_education_table.csv") # Save as csv

ldl_cohort_education_power_med_table <- power_ldl_cohort_education$power_median_dat     # Median power
ldl_cohort_education_power_med_table
write.csv (hdl_cohort_education_power_med_table, "ldl_cohort_education_power_med_table.csv") # Save as csv

firepower (list(power_ldl_cohort_education$power_median_dat))  # Firepower plot


#BMI cohort studies and education
View (databmi_cohort_education_pa)

power_bmi_cohort_education <- mapower_se(dat=databmi_cohort_education_pa,  # The dataset
                                     observed_es = 0.37,                # The observed summary effect size 
                                     name = "bmi_cohort_education_ma")  # The name of the meta-analysis

bmi_cohort_education_table <- power_bmi_cohort_education$dat  # Power for a range of effects
bmi_cohort_education_table
write.csv (bmi_cohort_education_table, "bmi_cohort_education_table.csv") # Save as csv

bmi_cohort_education_power_med_table <- power_bmi_cohort_education$power_median_dat     # Median power
bmi_cohort_education_power_med_table
write.csv (bmi_cohort_education_power_med_table, "bmi_cohort_education_power_med_table.csv") # Save as csv

firepower (list(power_bmi_cohort_education$power_median_dat))  # Firepower plot



#Allostatic load cohort studies and education
View (dataal_cohort_education_pa)

power_al_cohort_education <- mapower_se(dat=dataal_cohort_education_pa,  # The dataset
                                     observed_es = 0.45,                # The observed summary effect size 
                                     name = "al_cohort_education_ma")  # The name of the meta-analysis

al_cohort_education_table <- power_al_cohort_education$dat  # Power for a range of effects
al_cohort_education_table
write.csv (al_cohort_education_table, "al_cohort_education_table.csv") # Save as csv

al_cohort_education_power_med_table <- power_al_cohort_education$power_median_dat     # Median power
al_cohort_education_power_med_table
write.csv (al_cohort_education_power_med_table, "al_cohort_education_power_med_table.csv") # Save as csv

firepower (list(power_al_cohort_education$power_median_dat))  # Firepower plot

# Combining multiples meta-analysis
list_power <- list(al_cohort_education_power_med_table, crp_cohort_education_power_med_table, il6_cohort_education_power_med_table, tchol_cohort_education_power_med_table, crp_cohort_occupation_power_med_table, dbp_cohort_education_power_med_table, sbp_cohort_education_power_med_table, hdl_cohort_education_power_med_table, ldl_cohort_education_power_med_table, tri_cohort_education_power_med_table, bmi_cohort_education_power_med_table, wc_cohort_education_power_med_table)
firepower (list_power)


#ALL studies 
View (dataparental_sep_pa)

power_parental_sep <- mapower_se(dat=dataparental_sep_pa,  # The dataset
                                     observed_es = 0.16,                # The observed summary effect size 
                                     name = "parental_sep_ma")  # The name of the meta-analysis

parental_sep_pa_table <- power_parental_sep_pa$dat  # Power for a range of effects
parental_sep_pa_table
write.csv (parental_sep_pa_table, "parental_sep_pa_table.csv") # Save as csv

parental_sep_power_med_table <- power_parental_sep_pa$power_median_dat     # Median power
parental_sep_power_med_table
write.csv (parental_sep_power_med_table, "parental_sep_power_med_table.csv") # Save as csv

firepower (list(power_parental_sep$power_median_dat))  # Firepower plot


```

############################## END HERE ###########################






