---
title: "meta_analysis parental SEP and inflammation"
author: "Juan Rivillas"
date: "02/02/2023"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


#Install packages require to complete analysis
```{r}
install.packages("meta")
library(meta)
library(readxl)
```

***CRP***

#Upload table with studies CRP biomarkers 
```{r}
setwd("~/Documents/PhD Project/Literature review/systematic review 1 inflammation")

library(readxl)
datacrp <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/datacrp.xlsx")
View(datacrp)

View(datacrp)
head(datacrp)
```

#Check numerical values are not characters
```{r}
# datacrp$n_non_exposed     <- as.numeric(as.character(datacrp$n_non_exposed))
# datacrp$n_exposed         <- as.numeric(as.character(datacrp$n_exposed))
```


# Meta-analysis with continuout outcome
# comb.fixed/comb.random: indicator whether a fix/random effect mata-analysis to be conducted.
# sm: Three different types of summary measures to choose,standardized mean difference (SMD),mean difference (MD), ratio of means (ROM)
```{r}
res.crp =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = datacrp1, sm = "SMD")
res.crp
```


#Random forest regression to create an uncorrelated forest of decision trees.
#Random Forest is an ensemble of decision trees algorithms that can be used for classification and regression predictive modeling.
```{r}
forest(res.crp, leftcols = c('studlab', 'Study_design', 'Year', 'Age', 'Sex', 'Exposure'))
forest.meta(res.crp, leftcols = c('studlab', 'Year', 'Exposure'), layout = "RevMan5")
forest.meta(res.crp, layout = "RevMan5")

```


#Funnel Plot to check or Test for funnel plot asymmetry, based on rank correlation or linear regression method (Metabias).
#Funnel plots are a visual tool for investigating publication and other bias in meta-analysis. They are simple scatterplots of the treatment effects estimatedfrom individual studies (horizontal axis) against a measure of study size (vertical axis).
```{r}
funnel(res.crp)
```


#Use Egger’s test to check publication bias, can take string ‘Egger’ or ‘linreg’.
#Egger’s test is a linear regression of the intervention effect estimates on their standard errors weighted by their inverse variance
```{r}
metabias(res.crp, method.bias = 'linreg', k.min = 7, plotit = T)
```

The p-value is 0.3843 which implies no publication bias. However, this meta-analysis contains k=7 studies. Egger’s test may lack the statistical power to detect bias when the number of studies is small (i.e., k<10).


****IL-6*****

library(readxl)
datail6 <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/datail6.xlsx")
View(datail6)

```{r}
# meta-analysis with continuout outcome
# comb.fixed/comb.random: indicator whether a fix/random effect mata-analysis to be conducted.
# sm: Three different types of summary measures to choose,standardized mean difference (SMD),mean difference (MD), ratio of means (ROM)
res.il6 =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = datail6, sm = "SMD")
res.il6
```


```{r}
forest(res.il6, leftcols = c('studlab', 'Study_design', 'Year', 'Age', 'Sex', 'Exposure'))
forest.meta(res.il6, leftcols = c('studlab', 'Year', 'Exposure'), layout = "RevMan5")
forest.meta(res.il6, layout = "RevMan5")
```


```{r}
funnel(res.il6)
```

Use Egger’s test to check publication bias, can take string ‘Egger’ or ‘linreg’.
```{r}
metabias(res.il6, method.bias = 'linreg', k.min = 5, plotit = T)
```


***Allostatic Load***

> library(readxl)
> dataal <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/dataal.xlsx")
> View(dataal)  

```{r}
#dataal$mean_non_exposed     <- as.numeric(as.character(dataal$mean_non_exposed))
#dataal$mean_exposed         <- as.numeric(as.character(dataal$mean_exposed))
```


```{r}
# meta-analysis with continuout outcome
# comb.fixed/comb.random: indicator whether a fix/random effect mata-analysis to be conducted.
# sm: Three different types of summary measures to choose,standardized mean difference (SMD),mean difference (MD), ratio of means (ROM)
res.al =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = dataal, sm = "SMD")
res.al
```


```{r}
forest(res.al, leftcols = c('studlab', 'Study_design', 'Year', 'Age', 'Sex', 'Exposure'))
forest.meta(res.al, leftcols = c('studlab', 'Year', 'Exposure'), layout = "RevMan5")
forest.meta(res.al, layout = "RevMan5")
```


```{r}
funnel(res.al)
```


Use Egger’s test to check publication bias, can take string ‘Egger’ or ‘linreg’.
```{r}
metabias(res.al, method.bias = 'linreg', k.min = 3, plotit = T)
```



***DBP***
```{r}
library(readxl)
datasbp <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/datadbp.xlsx")
View(datadbp)
```


```{r}
# meta-analysis with continuout outcome
# comb.fixed/comb.random: indicator whether a fix/random effect mata-analysis to be conducted.
# sm: Three different types of summary measures to choose,standardized mean difference (SMD),mean difference (MD), ratio of means (ROM)
res.dbp =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = datadbp, sm = "SMD")
res.dbp
```

```{r}
forest(res.dbp, leftcols = c('studlab', 'Study_design', 'Year', 'Age', 'Sex', 'Exposure'))
forest.meta(res.dbp, leftcols = c('studlab', 'Year', 'Exposure'), layout = "RevMan5")
forest.meta(res.dbp, layout = "RevMan5")
```


```{r}
funnel(res.dbp)
```



Use Egger’s test to check publication bias, can take string ‘Egger’ or ‘linreg’.
```{r}
metabias(res.dbp, method.bias = 'linreg', k.min = 4, plotit = T)
```


#####RUN UNTIL HERE######

Total cholesterol

```{r}
library(readxl)
datatchol <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/datatchol.xlsx")
View(datatchol)
```


```{r}
# meta-analysis with continuout outcome
# comb.fixed/comb.random: indicator whether a fix/random effect mata-analysis to be conducted.
# sm: Three different types of summary measures to choose,standardized mean difference (SMD),mean difference (MD), ratio of means (ROM)
res.tchol =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = datatchol, sm = "SMD")
res.tchol
```

```{r}
forest(res.tchol, leftcols = c('studlab', 'Study_design', 'Year', 'Age', 'Sex', 'Exposure'))
```

```{r}
funnel(res.tchol)
```

LDL
```{r}
library(readxl)
dataldl <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/dataldl.xlsx")
View(dataldl)
```


```{r}
# meta-analysis with continuout outcome
# comb.fixed/comb.random: indicator whether a fix/random effect mata-analysis to be conducted.
# sm: Three different types of summary measures to choose,standardized mean difference (SMD),mean difference (MD), ratio of means (ROM)
res.ldl =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = dataldl, sm = "SMD")
res.ldl
```


```{r}
forest(res.ldl, leftcols = c('studlab', 'Study_design', 'Year', 'Age', 'Sex', 'Exposure'))
```


```{r}
funnel(res.ldl)
```



HDL
library(readxl)
datahdl <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/datahdl.xlsx")
View(datahdl)

```{r}
# meta-analysis with continuout outcome
# comb.fixed/comb.random: indicator whether a fix/random effect mata-analysis to be conducted.
# sm: Three different types of summary measures to choose,standardized mean difference (SMD),mean difference (MD), ratio of means (ROM)
res.hdl =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = datahdl, sm = "SMD")
res.ldl
```


```{r}
forest(res.hdl, leftcols = c('studlab', 'Study_design', 'Year', 'Age', 'Sex', 'Exposure'))
```


```{r}
funnel(res.hdl)
```

WC
library(readxl)
datawc <- read_excel("Documents/PhD Project/Literature review/systematic review 1 inflammation/datawc.xlsx")
View(datawc)

```{r}
# meta-analysis with continuout outcome
# comb.fixed/comb.random: indicator whether a fix/random effect mata-analysis to be conducted.
# sm: Three different types of summary measures to choose,standardized mean difference (SMD),mean difference (MD), ratio of means (ROM)
res.wc =  metacont(n_exposed, mean_exposed, sd_exposed,
n_non_exposed, mean_non_exposed, sd_non_exposed,
comb.fixed = T, comb.random = T, studlab = Study,
data = datawc, sm = "SMD")
res.wc
```


```{r}
forest(res.wc, leftcols = c('studlab', 'Study_design', 'Year', 'Age', 'Sex', 'Exposure'))
```

```{r}
funnel(res.wc)
```







